---
title: "openEvalR"
author: "Alina Hagen"
date: "`r Sys.Date()`"
output: html_document
---
``` {r}
# Necessary Libraries
library(jsonlite)
library(tidyverse)
library(irr)
library(caret)

```

# Results Converter:
This function will read a JSONL file containing model-generated outputs, extract relevant fields, and reshape the data so that outputs from multiple models for each unique item are displayed side by side in a standard dataframe. 

Necessary inputs: the OpenAI .jsonl Evaluation file path, the full name of the ground-truth column
Expected output: A dataframe consisiting of the data_source_idx, the ground-truth column, and any model-result pairing for each model in the test.

```{r}
results <- function(jsonl_path, groundtruth_col) {
  # Load JSONL file
  json_data <- stream_in(file(jsonl_path), flatten = TRUE)

  # Extract relevant fields and flatten model outputs
  json <- json_data %>%
    mutate(model_output = map_chr(sample.outputs, ~ .x[[2]])) %>%
    select(data_source_idx, !!sym(groundtruth_col), sample.sampled_model_name, model_output)

  # Assign model index to each sample
  json_prepped <- json %>%
    group_by(data_source_idx) %>%
    arrange(sample.sampled_model_name, .by_group = TRUE) %>%
    mutate(model_index = row_number()) %>%
    ungroup()

  # Pivot to wide format
  json_wide <- json_prepped %>%
    pivot_wider(
      id_cols = c(data_source_idx, !!sym(groundtruth_col)),
      names_from = model_index,
      values_from = c(sample.sampled_model_name, model_output),
      names_glue = "model_{model_index}_{.value}"
    )

  # Reorder model columns: name then output per model
  all_cols <- names(json_wide)
  meta_cols <- all_cols[1:2]  # assumes data_source_idx + groundtruth
  model_cols <- setdiff(all_cols, meta_cols)

  # Extract numeric order while preserving order
  model_indices <- unique(str_extract(model_cols, "(?<=model_)\\d+"))

  # Interleave model name + output per index
  reordered_model_cols <- unlist(lapply(model_indices, function(i) {
    grep(paste0("^model_", i, "_(sample\\.sampled_model_name|model_output)$"), model_cols, value = TRUE)
  }))

  # Recombine and return
  json_wide <- json_wide %>% select(all_of(c(meta_cols, reordered_model_cols)))
  return(json_wide)
}
```

# results() Example
```{r}
df <- results("Luigi_Raw.jsonl", "item.Luigi_stance")
head(df)

```

# Metadata Extractor
This function extracts item.* fields and data_source_idx from an OpenAI evaluation export, allowing users to reconnect model results to their original dataset using the unique identifier extracted using the results() function.

Necessary inputs: the OpenAI .jsonl Evaluation file path
Expected output: A dataframe consisiting of the data_source_idx and the original dataset uploaded to OpenAI

```{r}

metadata <- function(jsonl_path) {
  # Load the JSONL file
  json_data <- stream_in(file(jsonl_path), flatten = TRUE)

  # Grab all item.* columns
  item_cols <- grep("^item\\.", names(json_data), value = TRUE)

  # Add data_source_idx explicitly
  result <- json_data %>%
    select(data_source_idx, all_of(item_cols)) %>%
    arrange(data_source_idx)

  return(result)
}

```

# metadata() Example
``` {r}
df2 <- metadata("Luigi_Raw.jsonl")
head(df)

```

# Kappa Calculator 
This function calculates Cohen’s Kappa between a specified ground truth column and each model's output column in a dataframe. It helps evaluate how closely each model agrees with the human-labeled data, providing a standardized metric of inter-annotator reliability.

Necessary inputs: the dataframe created by the results() function, the full name of the ground-truth column
Expected output: a dataframe consisting of the Cohen's kappa between each model run and the ground truth

``` {r}
kappas <- function(df, groundtruth_col) {
  # Get all model output columns (assumes naming convention used by `converter()`)
  model_output_cols <- grep("model_\\d+_model_output", names(df), value = TRUE)
  
  # Calculate Cohen’s Kappa for each model
  results <- map_dfr(model_output_cols, function(model_col) {
    data <- df %>%
      select(!!sym(groundtruth_col), !!sym(model_col)) %>%
      drop_na()

    # Ensure both columns are factors for proper kappa handling
    data <- data %>% mutate_all(as.factor)

    kappa_score <- kappa2(data)$value

    tibble(
      model_output = model_col,
      groundtruth = groundtruth_col,
      kappa = kappa_score
    )
  })

  return(results)
}

```

# kappas() Example
``` {r}
df3 <- kappas(df, "item.Luigi_stance")
head(df3)

```

# Confusion Matrix
This function computes a confusion matrix between a specified ground truth column and each model's output column in a dataframe.

Necessary inputs: the dataframe created by the results() function, the full name of the ground-truth column
Expected output: a named list of caret::confusionMatrix objects for performance comparison.
```{r}
cmatrix <- function(df, groundtruth_col) {
  # Find all model output columns
  model_output_cols <- grep("model_\\d+_model_output", names(df), value = TRUE)
  
  # Create confusion matrix for each model
  confusion_list <- map(model_output_cols, function(model_col) {
    data <- df %>%
      select(!!sym(groundtruth_col), !!sym(model_col)) %>%
      drop_na() %>%
      mutate(across(everything(), as.factor))  # Ensure factors

    cm <- confusionMatrix(data[[model_col]], data[[groundtruth_col]])
    return(cm)
  })

  names(confusion_list) <- model_output_cols
  return(confusion_list)
}

```

# cmatrix() Example
```{r}
df4 <- cmatrix(df, "item.Luigi_stance")
print(df4)
```
